### 2020-03-25 Paper_shortened_pdf.Rmd
Editor comment 6

We hypothesized that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment. Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment. This means that, in families where parents displayed more consistent parenting, children had better emotional adjustment - and in families where parents engaged in more inductive reasoning or anger, children had poorer emotional adjustment.

@@

We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)). These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in children's emotional adjustment. In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment. This means that if, in one particular wave, parents displayed more warmth than usual, their children tended to display better emotional adjustment than usual. Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4). This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.

@@

In our analyses, each family has family-specific regression coefficients for the cross-lagged effects between wave-to-wave fluctuations in qualitative parenting, and emotional adjustment. Our hypotheses $H_{1a-d}$ also apply to these within-family predictive effects. To test these hypotheses, we examine the average standardized regression coefficients across families (see Table \@ref(tab:tabcrosspath)). In technical terms, these are random intercepts of the within-family predictive effects. Their variances indicate whether the family-specific regression coefficient differs in strength (and sign) across families. With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave. The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave. We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects. The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

@@

To examine interactions between involvement quality and quantity, we investigated whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, a one-standard deviation increase in joint involvement predicted a $-.14$ standard deviation decrease in the within-families predictive effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)). This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that parents' greater joint quantitative involvement, and mothers' greater relative involvement, both negatively predicted the within-family effect of adjustment on maternal warmth. We found no effects for consistency. For inductive reasoning, we found that greater joint involvement negatively predicted the within-family effect of adjustment on maternal inductive reasoning. Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

>>>>

We hypothesized that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment. At the between-family level, we examined correlations between families' average levels of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment. Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment. This means that, in families where parents reported more consistency, children displayed better emotional adjustment - and in families where parents reported more inductive reasoning or anger, children showed poorer emotional adjustment.

@@

We also examined within-family correlations, which are based on within-family fluctuations in the study variables (Table \@ref(tab:tabcors)). These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in their children's emotional adjustment. In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment. This means that if, in any particular wave, parents report more warmth than usual, their children displayed better emotional adjustment than usual. Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4). This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.

@@

In addition to the aforementioned within-family correlations, we also estimated over-time predictive effects between within-family fluctuations in qualitative parenting, and emotional adjustment. Each family has family-specific regression coefficients, and across families, we estimate the intercept and variance of these cross-lagged effects. The intercept indicates the average of the within-family predictive effects across families, and the variance indicate whether the family-specific regression coefficient varies in strength (and sign) across families. Hypotheses $H_{1a-d}$ also apply to these within-family predictive effects. To test these hypotheses, we examine the average standardized regression coefficients (see Table \@ref(tab:tabcrosspath)). With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave. The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave. We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects. The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

@@

We conducted exploratory analyses of interactions between involvement quality and quantity, by investigating whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). The novelty of this research area precluded us from formulating specific hypotheses. With regard to parenting effects, the only significant moderation we found was for maternal warmth. Specifically, in families where both parents were more involved, the cross-lagged effect of maternal warmth on emotional adjustment was *lower* than in families with less-involved parents. Specifically, $+1SD$ in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)). This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that greater joint quantitative involvement, and mothers' greater relative involvement, both negatively moderated the within-family effect of emotional adjustment on maternal warmth. We found no effects for consistency. For inductive reasoning, we found that greater joint involvement negatively moderated the within-family effect of adjustment on maternal inductive reasoning. Greater relative involvement also negatively moderated the effect of emotional adjustment on fathers' inductive reasoning. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment predicted a stronger decrease in paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.
